```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r pressure, echo=FALSE}
rm(list = ls())
gc()
```


```{r}
# Include Foreign Package To Read SAS Transport Files
library(foreign)
library(dummies)
library(tidyr)
library(dplyr)
```

```{r}
# Create Temporary File To Store Downloaded SAS Transport Files
tf <- tempfile()
```

#2015-2016
#Demographic Data 

```{r}
# Download NHANES 2015-2016 Demographic Data 
download.file("https://wwwn.cdc.gov/Nchs/Nhanes/2015-2016/DEMO_I.XPT", tf, mode="wb")
# Create Data Frame From Temporary File & Keep Variables Of Interest                                              
DEMO <- read.xport(tf)[,c("SEQN","RIAGENDR","RIDAGEYR","RIDRETH3")]

DEMO$RIAGENDR <- case_when(DEMO$RIAGENDR==1~1, DEMO$RIAGENDR==2~0)
DEMO$RIAGENDR <- as.factor(DEMO$RIAGENDR)

DEMO$Race <- factor(DEMO$RIDRETH3, labels=c("Mexican American","Other Hispanic","Non-Hispanic White","Non-Hispanic Black","6	Non-Hispanic Asian","Other Race - Including Multi-Racial"))
Race <- dummy(DEMO$Race)
demo <-as.data.frame(cbind(SEQN=DEMO$SEQN,Gender=DEMO$RIAGENDR,Age=DEMO$RIDAGEYR,Race))
demo$Gender <- case_when(demo$Gender==1~1, demo$Gender==2~0)
demo$Gender <- as.factor(demo$Gender)
nrow(demo)
```




#Diabetes, heart disease, family history, health condition

```{r}
# Download NHANES 2015-2016 Medical Conditions Data 
download.file("https://wwwn.cdc.gov/Nchs/Nhanes/2015-2016/MCQ_I.XPT", tf, mode="wb")
# Create Data Frame From Temporary File & Keep Variables Of Interest                                              
medic <- read.xport(tf)[,c("SEQN","MCQ300A","MCQ300B","MCQ300C","MCQ365A","MCQ365B","MCQ365C","MCQ365D","MCQ160K","MCQ230A")]
medic[is.na(medic)]<-0

medic$MCQ300A = case_when(medic$MCQ300A == 1 ~ 1, medic$MCQ300A == 2 ~ 0)
medic$MCQ300A = as.factor(medic$MCQ300A)

medic$MCQ300B = case_when(medic$MCQ300B == 1 ~ 1, medic$MCQ300B == 2 ~ 0)
medic$MCQ300B = as.factor(medic$MCQ300B)

medic$MCQ300C = case_when(medic$MCQ300C == 1 ~ 1, medic$MCQ300C == 2 ~ 0)
medic$MCQ300C = as.factor(medic$MCQ300C)

medic$MCQ365A = case_when(medic$MCQ365A == 1 ~ 1, medic$MCQ365A == 2 ~ 0)
medic$MCQ365A = as.factor(medic$MCQ365A)

medic$MCQ365B = case_when(medic$MCQ365B == 1 ~ 1, medic$MCQ365B == 2 ~ 0)
medic$MCQ365B = as.factor(medic$MCQ365B)


medic$MCQ365C = case_when(medic$MCQ365C == 1 ~ 1, medic$MCQ365C == 2 ~ 0)
medic$MCQ365C = as.factor(medic$MCQ365C)

medic$MCQ365D = case_when(medic$MCQ365D == 1 ~ 1, medic$MCQ365D == 2 ~ 0)
medic$MCQ365D = as.factor(medic$MCQ365D)

medic$MCQ160K = case_when(medic$MCQ160K == 1 ~ 1, medic$MCQ160K == 2 ~ 0)
medic$MCQ160K = as.factor(medic$MCQ160K)

med <- medic %>% select(-"MCQ230A") %>% drop_na()

```


#Smoke

```{r}
# Download NHANES 2015-2016 Smoke Data 
download.file("https://wwwn.cdc.gov/Nchs/Nhanes/2015-2016/SMQ_I.XPT", tf, mode="wb")
# Create Data Frame From Temporary File & Keep Variables Of Interest                                              
smoke<- read.xport(tf)[,c("SEQN","SMQ020","SMQ890","SMQ900","SMQ910")]
smoke$SMQ020<- case_when(smoke$SMQ020==1 ~ 1, smoke$SMQ020==2 ~ 0)
smoke$SMQ020<- as.factor(smoke$SMQ020)
smoke$SMQ890<- case_when(smoke$SMQ890==1 ~ 1, smoke$SMQ890==2 ~ 0)
smoke$SMQ890<- as.factor(smoke$SMQ890)
smoke$SMQ900<- case_when(smoke$SMQ900==1 ~ 1, smoke$SMQ900==2 ~ 0)
smoke$SMQ900<- as.factor(smoke$SMQ900)
smoke$SMQ910<- case_when(smoke$SMQ910==1 ~ 1, smoke$SMQ910==2 ~ 0)
smoke$SMQ910<- as.factor(smoke$SMQ910)

smoke<- na.omit(smoke)

```

#Exposure to second-hand smoke

```{r}
# Download NHANES 2015-2016 Exposure to second-hand smoke Data 
download.file("https://wwwn.cdc.gov/Nchs/Nhanes/2015-2016/SMQFAM_I.XPT", tf, mode="wb")
# Create Data Frame From Temporary File & Keep Variables Of Interest                                              
secondsmoke<- read.xport(tf)[,c("SEQN","SMD460")]
secondsmoke<-secondsmoke[!(secondsmoke$SMD460==777 | secondsmoke$SMD460==999),]

```
#Chest Pain 

```{r}
# Download NHANES 2015-2016 Chest Pain  Data 
download.file("https://wwwn.cdc.gov/Nchs/Nhanes/2015-2016/CDQ_I.XPT", tf, mode="wb")
# Create Data Frame From Temporary File & Keep Variables Of Interest                                              
chestpain<- read.xport(tf)[,c("SEQN","CDQ001")]
chestpain$CDQ001<- case_when(chestpain$CDQ001==1 ~ 1, chestpain$CDQ001==2 ~ 0)
chestpain$CDQ001<- as.factor(chestpain$CDQ001)
chestpain<- na.omit(chestpain)
```

#Volatile Toxicant

```{r}
# Download NHANES 2015-2016 Volatile Toxicant  Data 
download.file("https://wwwn.cdc.gov/Nchs/Nhanes/2015-2016/VTQ_I.XPT", tf, mode="wb")
# Create Data Frame From Temporary File & Keep Variables Of Interest                                              
Toxicant<- read.xport(tf)[,c("SEQN","VTQ281A","VTQ281C","VTQ281E")]

Toxicant$VTQ281A <- case_when(Toxicant$VTQ281A == 1 ~ 1, Toxicant$VTQ281A == 2 ~ 0)  
Toxicant$VTQ281A <- as.factor(Toxicant$VTQ281A)
Toxicant$VTQ281C <- case_when(Toxicant$VTQ281C == 1 ~ 1, Toxicant$VTQ281C == 2 ~ 0)  
Toxicant$VTQ281C <- as.factor(Toxicant$VTQ281C)
Toxicant$VTQ281E <- case_when(Toxicant$VTQ281E == 1 ~ 1, Toxicant$VTQ281E == 2 ~ 0)  
Toxicant$VTQ281E <- as.factor(Toxicant$VTQ281E)
Toxicant <- na.omit(Toxicant)
```

#Arsenic - Total - Urine
```{r}
# Download NHANES 2015-2016 Arsenic  Data 
download.file("https://wwwn.cdc.gov/Nchs/Nhanes/2015-2016/UTAS_I.XPT", tf, mode="wb")
# Create Data Frame From Temporary File & Keep Variables Of Interest                                              
Arsenic<- read.xport(tf)[,c("SEQN","URXUAS")]
Arsenic[is.na(Arsenic)]<-0
```


#LBDBCRSI - Chromium (nmol/L)
```{r}
# Download NHANES 2015-2016 Chromium  Data 
download.file("https://wwwn.cdc.gov/Nchs/Nhanes/2015-2016/CRCO_I.XPT", tf, mode="wb")
# Create Data Frame From Temporary File & Keep Variables Of Interest                                              
Chromium<- read.xport(tf)[,c("SEQN","LBDBCRSI")]
Chromium[is.na(Chromium)]<-0
```

#DR1TBCAR - Beta-carotene (mcg)
```{r}
# Download NHANES 2015-2016 Beta-carotene Data 
download.file("https://wwwn.cdc.gov/Nchs/Nhanes/2015-2016/DR1TOT_I.XPT", tf, mode="wb")
# Create Data Frame From Temporary File & Keep Variables Of Interest                                              
Betacarotene1<- read.xport(tf)[,c("SEQN","DR1TBCAR")]%>% drop_na()
 
download.file("https://wwwn.cdc.gov/Nchs/Nhanes/2015-2016/DR2TOT_I.XPT", tf, mode="wb")
# Create Data Frame From Temporary File & Keep Variables Of Interest                                              
Betacarotene2<- read.xport(tf)[,c("SEQN","DR2TBCAR")]%>% drop_na()
Betacarotene <- merge(Betacarotene1,Betacarotene2, by = "SEQN") 
Betacarotene$avg_betacarotene <- apply(Betacarotene[,2:3], 1, mean)
Betacarotene <- Betacarotene %>% select("SEQN", "avg_betacarotene")
```


# Merge all tables
```{r}
alltable<-Reduce(function(x,y) merge(x = x, y = y, by = "SEQN"), 
                 list(demo,med, smoke, secondsmoke,chestpain,Toxicant, Arsenic,Chromium,Betacarotene))

```

# Tareget
```{r}
lung <- medic[medic$MCQ230A==23,] # have lung cancer

Records_lung <-lung %>% mutate(y = 1) 
Records_lung<-Records_lung[c("SEQN","y")]
nrow(Records_lung)

Healthy_Records <- alltable[-Records_lung$SEQN,] %>% mutate(y = 0)
Healthy_Records <-Healthy_Records[c("SEQN","y")]
table = rbind(Healthy_Records,Records_lung)
table<-table[c("SEQN","y")]

table_final = merge(alltable,table, by = "SEQN")
table_final[is.na(table_final)]<-0
nrow(table_final)

```






```{r}

# normalize input variables
colnames(table_final)

selected<- table_final[,c(3,22,27:29)]
unse<-table_final[,-c(3,22,27:29)]
normalization<-function(x){
  return((x-min(x))/(max(x)-min(x)))}
norm<-normalization(selected)

table_final1<-cbind(unse[,1:24],norm,y=unse$y)

write.csv(table_final1,"/Users/tmh/Desktop/MedBlob/lung cancer/lungcancer(normalized).csv")


```

#2017-2018
#Demographic Data 

```{r}
# Download NHANES 2017-2018 Demographic Data 
download.file("https://wwwn.cdc.gov/Nchs/Nhanes/2017-2018/DEMO_J.XPT", tf, mode="wb")
# Create Data Frame From Temporary File & Keep Variables Of Interest                                              
DEMO1 <- read.xport(tf)[,c("SEQN","RIAGENDR","RIDAGEYR","RIDRETH3")]

DEMO1$RIAGENDR <- case_when(DEMO1$RIAGENDR==1~1, DEMO1$RIAGENDR==2~0)
DEMO1$RIAGENDR <- as.factor(DEMO1$RIAGENDR)

DEMO1$Race <- factor(DEMO1$RIDRETH3, labels=c("Mexican American","Other Hispanic","Non-Hispanic White","Non-Hispanic Black","6	Non-Hispanic Asian","Other Race - Including Multi-Racial"))
Race1 <- dummy(DEMO$Race)
demo <-as.data.frame(cbind(SEQN=DEMO$SEQN,Gender=DEMO$RIAGENDR,Age=DEMO$RIDAGEYR,Race))
demo$Gender <- case_when(demo$Gender==1~1, demo$Gender==2~0)
demo$Gender <- as.factor(demo$Gender)
nrow(demo)
```




#Chronic bronchitis, family history, health condition

```{r}
# Download NHANES 2017-2018 Medical Conditions Data 
download.file("https://wwwn.cdc.gov/Nchs/Nhanes/2017-2018/MCQ_J.XPT", tf, mode="wb")
# Create Data Frame From Temporary File & Keep Variables Of Interest                                              
medic <- read.xport(tf)[,c("SEQN","MCQ300A","MCQ300B","MCQ300C","MCQ366A","MCQ366B","MCQ366C","MCQ366D","MCQ160K","MCQ230A")]
medic[is.na(medic)]<-0

medic$MCQ300A = case_when(medic$MCQ300A == 1 ~ 1, medic$MCQ300A == 2 ~ 0)
medic$MCQ300A = as.factor(medic$MCQ300A)

medic$MCQ300B = case_when(medic$MCQ300B == 1 ~ 1, medic$MCQ300B == 2 ~ 0)
medic$MCQ300B = as.factor(medic$MCQ300B)

medic$MCQ300C = case_when(medic$MCQ300C == 1 ~ 1, medic$MCQ300C == 2 ~ 0)
medic$MCQ300C = as.factor(medic$MCQ300C)

medic$MCQ366A = case_when(medic$MCQ366A == 1 ~ 1, medic$MCQ366A == 2 ~ 0)
medic$MCQ366A = as.factor(medic$MCQ366A)

medic$MCQ366B = case_when(medic$MCQ366B == 1 ~ 1, medic$MCQ366B == 2 ~ 0)
medic$MCQ366B = as.factor(medic$MCQ366B)


medic$MCQ366C = case_when(medic$MCQ366C == 1 ~ 1, medic$MCQ366C == 2 ~ 0)
medic$MCQ366C = as.factor(medic$MCQ366C)

medic$MCQ366D = case_when(medic$MCQ366D == 1 ~ 1, medic$MCQ366D == 2 ~ 0)
medic$MCQ366D = as.factor(medic$MCQ366D)

medic$MCQ160K = case_when(medic$MCQ160K == 1 ~ 1, medic$MCQ160K == 2 ~ 0)
medic$MCQ160K = as.factor(medic$MCQ160K)

med <- medic %>% select(-"MCQ230A") %>% drop_na()

```


#Smoke

```{r}
# Download NHANES 2017-2018 Smoke Data 
download.file("https://wwwn.cdc.gov/Nchs/Nhanes//SMQ_J.XPT", tf, mode="wb")
# Create Data Frame From Temporary File & Keep Variables Of Interest                                              
smoke<- read.xport(tf)[,c("SEQN","SMQ020","SMQ890","SMQ900","SMQ910")]
smoke$SMQ020<- case_when(smoke$SMQ020==1 ~ 1, smoke$SMQ020==2 ~ 0)
smoke$SMQ020<- as.factor(smoke$SMQ020)
smoke$SMQ890<- case_when(smoke$SMQ890==1 ~ 1, smoke$SMQ890==2 ~ 0)
smoke$SMQ890<- as.factor(smoke$SMQ890)
smoke$SMQ900<- case_when(smoke$SMQ900==1 ~ 1, smoke$SMQ900==2 ~ 0)
smoke$SMQ900<- as.factor(smoke$SMQ900)
smoke$SMQ910<- case_when(smoke$SMQ910==1 ~ 1, smoke$SMQ910==2 ~ 0)
smoke$SMQ910<- as.factor(smoke$SMQ910)

smoke<- na.omit(smoke)

```

#Exposure to second-hand smoke

```{r}
# Download NHANES 2015-2016 Exposure to second-hand smoke Data 
download.file("https://wwwn.cdc.gov/Nchs/Nhanes/2015-2016/SMQFAM_I.XPT", tf, mode="wb")
# Create Data Frame From Temporary File & Keep Variables Of Interest                                              
secondsmoke<- read.xport(tf)[,c("SEQN","SMD460")]
secondsmoke<-secondsmoke[!(secondsmoke$SMD460==777 | secondsmoke$SMD460==999),]

```
#Chest Pain 

```{r}
# Download NHANES 2015-2016 Chest Pain  Data 
download.file("https://wwwn.cdc.gov/Nchs/Nhanes/2015-2016/CDQ_I.XPT", tf, mode="wb")
# Create Data Frame From Temporary File & Keep Variables Of Interest                                              
chestpain<- read.xport(tf)[,c("SEQN","CDQ001")]
chestpain$CDQ001<- case_when(chestpain$CDQ001==1 ~ 1, chestpain$CDQ001==2 ~ 0)
chestpain$CDQ001<- as.factor(chestpain$CDQ001)
chestpain<- na.omit(chestpain)
```

#Volatile Toxicant

```{r}
# Download NHANES 2015-2016 Volatile Toxicant  Data 
download.file("https://wwwn.cdc.gov/Nchs/Nhanes/2015-2016/VTQ_I.XPT", tf, mode="wb")
# Create Data Frame From Temporary File & Keep Variables Of Interest                                              
Toxicant<- read.xport(tf)[,c("SEQN","VTQ281A","VTQ281C","VTQ281E")]

Toxicant$VTQ281A <- case_when(Toxicant$VTQ281A == 1 ~ 1, Toxicant$VTQ281A == 2 ~ 0)  
Toxicant$VTQ281A <- as.factor(Toxicant$VTQ281A)
Toxicant$VTQ281C <- case_when(Toxicant$VTQ281C == 1 ~ 1, Toxicant$VTQ281C == 2 ~ 0)  
Toxicant$VTQ281C <- as.factor(Toxicant$VTQ281C)
Toxicant$VTQ281E <- case_when(Toxicant$VTQ281E == 1 ~ 1, Toxicant$VTQ281E == 2 ~ 0)  
Toxicant$VTQ281E <- as.factor(Toxicant$VTQ281E)
Toxicant <- na.omit(Toxicant)
```

#Arsenic - Total - Urine
```{r}
# Download NHANES 2015-2016 Arsenic  Data 
download.file("https://wwwn.cdc.gov/Nchs/Nhanes/2015-2016/UTAS_I.XPT", tf, mode="wb")
# Create Data Frame From Temporary File & Keep Variables Of Interest                                              
Arsenic<- read.xport(tf)[,c("SEQN","URXUAS")]
Arsenic[is.na(Arsenic)]<-0
```


#LBDBCRSI - Chromium (nmol/L)
```{r}
# Download NHANES 2015-2016 Chromium  Data 
download.file("https://wwwn.cdc.gov/Nchs/Nhanes/2015-2016/CRCO_I.XPT", tf, mode="wb")
# Create Data Frame From Temporary File & Keep Variables Of Interest                                              
Chromium<- read.xport(tf)[,c("SEQN","LBDBCRSI")]
Chromium[is.na(Chromium)]<-0
```

#DR1TBCAR - Beta-carotene (mcg)
```{r}
# Download NHANES 2015-2016 Beta-carotene Data 
download.file("https://wwwn.cdc.gov/Nchs/Nhanes/2015-2016/DR1TOT_I.XPT", tf, mode="wb")
# Create Data Frame From Temporary File & Keep Variables Of Interest                                              
Betacarotene1<- read.xport(tf)[,c("SEQN","DR1TBCAR")]%>% drop_na()
 
download.file("https://wwwn.cdc.gov/Nchs/Nhanes/2015-2016/DR2TOT_I.XPT", tf, mode="wb")
# Create Data Frame From Temporary File & Keep Variables Of Interest                                              
Betacarotene2<- read.xport(tf)[,c("SEQN","DR2TBCAR")]%>% drop_na()
Betacarotene <- merge(Betacarotene1,Betacarotene2, by = "SEQN") 
Betacarotene$avg_betacarotene <- apply(Betacarotene[,2:3], 1, mean)
Betacarotene <- Betacarotene %>% select("SEQN", "avg_betacarotene")
```



# Merge all tables
```{r}
alltable<-Reduce(function(x,y) merge(x = x, y = y, by = "SEQN"), 
                 list(demo,med, smoke, secondsmoke,chestpain,Toxicant, Arsenic,Chromium,Betacarotene))

```

# Tareget
```{r}
lung <- medic[medic$MCQ230A==23,] # have lung cancer

Records_lung <-lung %>% mutate(y = 1) 
Records_lung<-Records_lung[c("SEQN","y")]
nrow(Records_lung)

Healthy_Records <- alltable[-Records_lung$SEQN,] %>% mutate(y = 0)
Healthy_Records <-Healthy_Records[c("SEQN","y")]
table = rbind(Healthy_Records,Records_lung)
table<-table[c("SEQN","y")]

table_final = merge(alltable,table, by = "SEQN")
table_final[is.na(table_final)]<-0
nrow(table_final)

```
